---
title: "Cigarettes"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Cigarettes

The file "cigarettes_consumption.csv" contains information about cigarettes in 1000 regions in the world.

Specifically, the features are:

-   **cigarettes**: number of cigarettes packages sold for inhabitant in the considered region

-   **age**: average age of the population in the considered region

-   **income**: the average income of the population in the considered region

-   **population**: million of inhabitants in the considered region

-   **price**: average price of a pack of cigarettes

-   **tax**: tax amout for a pack of cigarettes

-   **ethnicity**: ethnicity of the major part of the inhabitants ("White","Causasian","Hispanic")

-   **country**: binary (1: USA, 0: Otherwise)

-   **treatment**: binary variable (1: the considered region has been subject of the cigarettes consumption reduction programme, 0: otherwise)

```{r}
data <- read.csv("../Datasets/cigarettes_consumption.csv", sep=";")
data$population_normalized=NULL
head(data)
```

## Question

Evaluate the impact of price on the average quantity of cigarettes sold per inhabitant.

## OLS regression

```{r}
data$population <- scale(data$population)
formula = cigarettes ~ age + income + population + price + ethnicity + country
ols_model_1 = lm(formula, data)
summary(ols_model_1)
```

The price doesn't seem significant due to the low p-value (there is no statistical support to say that the influence of price over the number of sold packages is not 0)

## IV regression

```{r}
library(ivreg)
library(rdd)
library(rdrobust)
mod_iv = ivreg(cigarettes ~ age + income + population + price + ethnicity + country | age + income + population + tax + ethnicity + country, data = data)
summary(mod_iv, diagnostics = TRUE)
```
